`r if(knitr:::is_latex_output()) '\\appendix'`

`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'` 

# Further resources

For information on using R markdown with LaTeX, Yihui Xie's [bookdown guide](https://bookdown.org/yihui/bookdown/) is fairly comprehensive.

If you want to integrate more TeX commands into your thesis, the [Comprehensive TeX Archive Network](https://www.ctan.org/) is where you will find documentation on all the TeX packages available.

Chester Ismay's thesisdown included a guide (I've written this present guide in place of that one) that you can find [here](https://raw.githubusercontent.com/ismayc/thesisdown_book/master/thesis.pdf). You may also find Rosanna van Hespen's [series of blogposts](https://rosannavanhespen.nl/category/rmarkdown/) on writing a thesis in R markdown useful, although Rosanna used limited LaTeX code and wrote it directly into the markdown files.

# The second appendix

This second appendix includes all of the R chunks of code that were hidden throughout the document (using the `include = FALSE` chunk tag) to help with readability and/or setup.

```{r ref.label='include_packages', eval = FALSE, echo = TRUE}
```

```{r ref.label='knit-diagram', eval = FALSE, echo = TRUE}
```
